Untwisting the Tetrahelix (v0.9)

### 2017-03-02 doc/TetrahelixUntwist.tex
significant improvements

\title{Untwisting the Tetrahelix (v0.8)}
\author{Robert L. Read \texttt{read.robert@gmail.com} \and
  Robert Gatliff \texttt{robert@toubat.org}
}

@@

This formula defines a helix, but it is not any of the apparent helices, or rail helices, of the
BC helix, but rather one that winds three times as rapidly through all
nodes. To a designer of tetrahelices, it is more natural to think of
the three helices which are visually apparent, that is, those three
which are closely approximated by the by the outer edges or rails of
the BC helix. We think of each of these three rails as being a different color, red, blue, or yellow.

@@

Note in Figure \ref{railanglefig} the $z$-axis travel for one rail edge is denoted by $d$. In \eqref{graycoxeter} and \label{eq:colored}, the variable
$h$ is used for one third of the distance we name $d$. We will later justify that $d = 3h$. In this paper we assume the length of a rail
is alwyas $1$ is a simplification, although we make the rail length a parameter in our JavaScript codein \texttt{tetrahelixmath.js} \cite{readtetrahelix}.

Since:
\[ \frac{2 \pi}{\rho_{bc}} \approx 10.16
\]
We can see that there are approximately $10.16$ red, blue or yellow tetrahedra on one rail in a single revolution.
The pitch of the Boerdijk--Coxeter helix of edge length $1$ is the length of three tetrahedra times this number:
\begin{align*}
  &= \frac{d_{bc} 2 \pi }{\rho_{bc}} \\  
%%  &= \frac{3 \cdot h_{bc} 2 \pi }{\rho_{bc}} \\
  &= \frac{3  \sqrt{\frac{2}{5}}  \pi}{\rho_{bc}} \\
  &\approx 9.64 \\
\end{align*}
The pitch is less than the number of tetrahedra because the tetrahedra
are not lined up perfectly.  It is a famous and interesting result
that the pitch is irrational, a BC helix never has two tetrahedra at
precisely the same orientation around the $z$-axis. However, this is
inconvenient to designers, who might prefer a rational pitch.
The idea of deveoping a rational period by arranging solid tetrahedra by relaxing the face-to-face matching
has been explored\cite{sadler2013periodic}. In this paper
we develop slightly irregular edge lengths that support, for example, a pitch of precisely 10
tetrahedra in one revolution which would allow an architect to design a
column having a basis and a capital in the same relation to the
tetrahedra they touch at the bottom and top of the column.

@@

If all three rails do not have the same pitch, there is an edge of
unbounded length, so it is not optimal. So we are justified in talking about the
\emph{pitch} of 
the tetrahelix as the pitch of its rail helices, even though there are
three such helices.

@@

Since the axes are parallel, we may define the \emph{inradius} of a tetrahelix to be the radius of the largest
``in cylinder'': a
cylinder parallel to this axis contained within circumscribing cylinder which is penetrated by no edge.
Define an \emph{minimax edge-length optimal tetrahelix} or just an
\emph{optimal tetrahelix} to be a tetrahelix for which there exists
no other tetrahelix of the same inradius and pitch with a lower maximum
difference between its edge lengths.

@@

Since all three rails have the same rail length, no matter how we
move the rails in the $xy$ plane if we shorten the $xy$ distance between
vertices we shorten the total distance.
Our tool for thinking about this is to project out the $z$ dimension to form
a two-dimensional figure of nodes and non-rail edges.
Consider the projection along the $z$ axis of all vertices and non-rail edges into the $xy$ plane, which will be
a figure of dots and connecting segments in the $xy$ plane. The convex
hull for any one helix projection will be a circle (if its pitch is
irrational) or an polygon if rational, or a point if the helix has
zero curvature. Each of these figures by definitions lies outside the
circle of radius $r$ in the $xy$ plane.

@@

Figure \ref{untouchingrailfig} depicts a projection of
depicting only the point $P$, the point $P'$ which is an adjacent vertex on the same rail,
and the intermediary points $R$ and $Q$ on the other two rails.

Supoose there is a rail $P$ which does not have a segment touching the incircle
at all in the projection, as depicted in 
Then a segment connecting the other two $Q$ and $R$ is a longest
rail, because any chord touching an incircle is longer than any which does not.
These two rails ($Q$ and $R$) can be moved closer to each other, and further away from $P$, to form a better minimax solution
by shortening the longest rail, until one of the segments from $P$ the previously
untouching rail touches one of the incircle. By induction then all rails have
at least one segement touch each vertex on the rail that is tangent to the
the inradius in an optimal tetrahelix.

@@

Theorem \ref{thm:cylinder} is not true when $\rho$ is an odd multiple of $\pi$.

Now that we have show that any optimal tetrahelix vertices that
are on helices of the same axes and pitch, we see that the vertices 
of any optimal tetrahelix will lie on a cylinder, or a circle when axis dimension
is projected out. Therefore is is reasonable to now speak of the \emph{radius}
of a tetrahelix as the radius of the cylinder, and the notion
of the projected incircle radius is now obsolete.

Now that we have coincident axes, the same pitch, we can go on to
 the hard proof about where they occur along the $z$-axis.

@@

 Note that from the point of view of a single edge, we are on a
 slanted cylinder, when $\rho \neq 0$.  This means for its point of
 view a cross section is an ellipse. So we have to be very careful in
 comparing lengths of edges relative to the tetrahedron, because a
 change in position along the edge changes the length of a line, but
 in a complicated way depending on where it is relative to the
 ellipse.

@@

\begin{theorem}
   \label{eventhirds}
   A tetrahelix of a given radius and axial distance $d$ in which all nodes are evenly spaced at $d/3$ intervals on the $z$ axis is optimal.
     Any one tetrahedron in a tetrahelix has $1$ rail edge, $2$ one-hop edges connected to the rail and $2$ two-hop edges connected to the rail.
  The edge opposite of the rail edge is a one-hop edge.
\end{theorem}

@@

    Any displacemnt along the $z$ axis of any rail increases the
    length of one or two two-hop edges and shortens the length of one
    or two one-hop edges.  This increases the minimax distance no
    matter what the rail edge length is, since all rail edges are the
    same length. Therefore, an evenly spaced tetrahelix is the unique
    optimal for any given radius and height.
\end{proof}

@@

This formulation $V(n,c)$ above is valuable, but obscures the essentially fact that the red, yellow, and blue helices distributed
about the central $z$ axis $120\degree$ from each other.
In order to rewrite this expression with an explicit rotation of $2\pi/3$, we expand 
the expression and seek to isolate the term $c2\pi/3$.
\begin{align*}
  \rho_{bc} n + c \theta  &=   \text{\{we aim for 3 in denominator, so we split...\}} \\
    (3 \theta - 2 \pi)n + (c/3)  (3 \theta)  &=   \text{\{we want $2\pi$ in numerator, so add canceling terms...\}} \\
  (3 \theta - 2 \pi)n + (c/ 3) (3 \theta - 2 \pi  + 2 \pi) &=  \text{\{associate...\}} \\
  (3 \theta - 2 \pi)n + (c/ 3) ((3 \theta - 2 \pi)  + 2 \pi) &=  \text{\{distribute...\}} \\  
  (3 \theta - 2 \pi)n + (c / 3) (3 \theta - 2 \pi)  + c 2 \pi /3 &=  \text{\{definition of $\rho_{bc}$...\}} \\
  \rho_{bc} n + (c / 3) \rho_{bc}  + c 2 \pi /3 &=  \text{\{collect like factors...\}} \\  
  \rho_{bc} (n + c/3)  + c 2 \pi /3  \\
\end{align*}
Now the the term on the left is the only one that depends on the scalar $n$. We use this to a create
a new formulation: $H_{BCsymmetric}(n,c) = H_{BCcolored}(n,c)$

The expression $n+c/3$ will now occur so often that we call it the ``c($\kappa$)olored number'' and we use the variable $\kappa$ to represent it: $\kappa = n+c/3$.
Recall that $c \in \{0,1,2\}$, but $n$ and $\kappa$ are are continuous (rational or real-valued.)

\begin{equation}
H_{BCsymmmetric}(n,c) =
\left [
  \begin{tabular}{c}
   $ r  \cos(\rho_{bc} \kappa  + c 2 \pi /3) $\\
   $ r  \sin(\rho_{bc} \kappa  + c 2 \pi /3) $\\
   $ \kappa 3  h_{bc} $
  \end{tabular}
  \right ],
\text{where:}
  \begin{tabular}{c}
 $\kappa = n + c/3$ \\
    $\rho_{bc} = (3 \theta - 2 \pi)$ \\
    $ \theta = \arccos(-2/3) $ \\
    $ h_{bc} = 1/\sqrt{10} $ \\    
  \end{tabular}      
\end{equation}

@@

Rewriting our formulation in terms of $\rho$:
\begin{equation}
  \label{eq:general}
H_{general}(\chi,n,c,\rho,d_{\rho},r_{\rho}) =
\left [
  \begin{tabular}{c}
   $ r_{\rho} \cos(\chi \cdot (\rho \kappa + c 2 \pi /3)) $\\
   $ r_{\rho}  \sin(\chi \cdot (\rho \kappa + c 2 \pi /3)) $\\
   $ d_{\rho} \kappa $
  \end{tabular}
  \right ],
\text{where:}
\begin{tabular}{c}
  $   1 = d_{\rho}^2 + 4 r_{\rho}^2 (\sin{ \rho / 2})^2 $ \\
  $\kappa = n + c/3$ \\
    $\chi \in \{-1,1\}$ \\  
  \end{tabular}      
\end{equation}

@@

\begin{align*}
  \text{one-hop} &= dist(H_{general}(n,c,\rho,d_{\rho}),H_{general}(n,c+1,\rho,d_{\rho}),r_{\rho})  \\
%%  \text{one-hop} &= dist(H_{general}(0,0,\rho,d_{\rho}),H_{general}(0,1,\rho,d_{\rho},r_{\rho}))  \\  
%%  \text{one-hop}  &= \sqrt{\frac{d_{\rho}^2}{9} + (r_{\rho}\sin(0) - r_{\rho}\sin(\rho/3+1\cdot\frac{2\pi}{3}))^2  +
%%    (r_{\rho}\cos(0) - r_{\rho}\cos(\rho/3 + 1\cdot\frac{2\pi}{3}))^2} \\
%%  \text{one-hop}  &= \sqrt{\frac{d_{\rho}^2}{9} + (0  - r_{\rho}\sin(\rho/3 + \frac{2\pi}{3}))^2  +
%%    (r_{\rho} - r_{\rho}\cos(\rho/3 + \frac{2\pi}{3}))^2} \\
%%  \text{one-hop}  &= \sqrt{\frac{d_{\rho}^2}{9} + r_{\rho}^2\sin^2(\rho/3 + \frac{2\pi}{3})  +
%%    r_{\rho}^2(1 - \cos(\rho/3 + \frac{2\pi}{3}))^2} \\
  %% Formula above checks at sqrt(8/27), 1, 0, and rho_bc, h_bc, r_bc.
  \text{one-hop}  &= \sqrt{\frac{d_{\rho}^2}{9} + r_{\rho}^2(\sin^2(\rho/3 + \frac{2\pi}{3})  + (1 - \cos(\rho/3 + \frac{2\pi}{3}))^2)} \\
  \text{but: }  d_{\rho}^2 &= 1 - 4 r_{\rho}^2 (\sin( \rho / 2)^2 \text{so we substitute:}\\
  %% Formula above checks at sqrt(8/27), 1, 0, and rho_bc, h_bc, r_bc.  
  \text{one-hop}  &= \sqrt{\frac{1}{9}  + r_{\rho}^2(-\frac{4 (\sin^2( \rho / 2))}{9} + \sin^2(\rho/3+ \frac{2\pi}{3})  + (1 - \cos(\rho/3 + \frac{2\pi}{3}))^2)} \\
  %% Formula above checks at sqrt(8/27), 1, 0, and rho_bc, h_bc, r_bc.    
\end{align*}

@@

\begin{theorem}
  \label{optimality}
  For any rail angle $\rho$ where $\rho \leq \rho_{bc}$, a tetrahelix of minimum maximum edge length
  difference is generated by $H_{general}(\chi,n,c,\rho,d_{\rho},r_{opt})$, where $r_{opt}$ and $d_{opt}$ are
  computed from $\rho$. 
\end{theorem}

\begin{proof}
  Theorem \ref{eventhirds} theorem tells us an optimal will have even thirds,
  as Equation \eqref{eq:general} does.
  Theorem \ref{generalformulaoptimal} tell us that
  that plugging $r_{opt}, d_{opt}$ into $H_{general}(\chi,n,c,\rho,d_{\rho},r_{\rho})$ minimizes the maximum length difference.
\end{proof}

If you look down the axis of an optimal tetrahelix, it happens that only the one-hop edges
come closest to the axis. In other words, they define the radius of the incircle of the
projection, or the radius of a cylinder that would just fit inside the tetrahelix.
This is useful to know if the tetrahelix is used as a structure to house a pipe or
a ladder that must bear a human or to armor a conduit. The \emph{inradius} of
an optimal tetrahelix is a remarkably simple function of the radius $r$ and the rail angle $\rho$:
\begin{equation}
  \label{eq:inradius}
  \frac{r \sin{(\pi - \rho)}}{6}
\end{equation}
Which can be seen from the trigonometry of the projection diagram:

@@

Choosing $d = 1$ and $\rho = 0$ we use Equation \eqref{eqrhoopt} to find the radius of 
optimal minimax difference.

@@

\begin{align*}
  \text{one-hop} &= \sqrt{\frac{1}{9} + 3r^2}\\
\end{align*}
Then:
\begin{align*}
   1  &=  \sqrt{\frac{1}{9} + 3r^2} \\
   r  &= \sqrt{\frac{8}{27}} \\
   r &\approx 0.5443
\end{align*}

This radius produces a two-hop rail length of
$\frac{2}{\sqrt{3}}$\footnote{Another interesting but non-optimal solution is derived by setting (one-hop + two-hop)/2 = 1,  occurs at $r = \sqrt{35}/4$,
which produces three length classes of $11/12, 12/12, 13/12$.}
. The difference between this 
and $1$ is $\approx 15.47\% $.

\begin{figure}[H]
  \label{fig:unfolded}
     \centering
     \includegraphics[width=0.6\textwidth]{figures/EquitetrabeamUnfolded.JPG}
     \caption{Projection Diagram (needs labels)}
\end{figure}

@@

 Because the equitetrabeam which as a rail angle of $0$ still has
 chirality, that is, one still must decide to connect with one-hop to
 the clockwise or counter-clockwise node, it is not poosible to build
 a smooth continuum where $\rho$ transitions from positive to negative
 which remains optimal. One can use a negative $\rho$ in $H_{general}$
 but it does not produce minimax optimal tetrahelices. In other words,
 untwisting a conter-clockwise tetrahelix to rail angle $0$ and then going
even further does produce a clockwise tetrahelix, but one in which the
 one-hop and two-hop lengths in the wrong places (that is, two-hop
 becomes shorter than one-hop.)

The pitch of any tetrahelix 
where $\rho \neq 0$ is:
\begin{equation}
  \label{pitcheqn}
p(\rho) = \frac{2 \pi  \cdot d}{\rho}
\end{equation}
For a fixed $z$-axis travel $d$, this is trivial. 
$z$-axis travel as per \eqref{dopt} pitch increases monotonically and smoothly with decreasing $\rho$, so
Equation \eqref{pitcheqn} can be easily solved numerically.
For a pitch at least $ p \geq \frac{3  \sqrt{\frac{2}{5}}  \pi}{\rho_{bc}} \approx 9.64 $,
using \eqref{dopt} produces minimax optimal tetrahelices.

In this way a rail angle can be chosen for any desired (possible) pitch, yield
the optimum radius, one-hop, and two-hop lengths an engineer needs to
construct a physical structure.

Perhaps surprisingly, the optimal untwisting is accomplished only by
changing the length of the two-hop member, leaving the one-hop memever
and rail length equivalent within this continuum.\footnote{Before deriving Equation \eqref{eqrhoopt}, we created a continuum by
using a linear interpolation between the optimal radius for the
Equitetrabeam and the BC Helix. This minimax optimum of this simpler
approach was at most 1\% worse than the optimum computed by
\eqref{eqrhoopt}.}
 However, it should
be noted that an engineer or architect may also use $H_{general}$
directly and interactively, and that minimax length optimality is a
mathematic notion only loosely related to the beauty and utility of
physical structures. For example, a structural engineer might increase
radius in order to resist buckling. If an equitetrabeam were actually
used as a beam, an engineer might start with the optimal tetrabeam and
dilate it in one dimension to ``deepen'' the beam. Similarly, simple
changes curve the equitetrabeam into an ``arch''.

@@

    \section{Old Proofs}

We can create the distance between representative nodes as:
\begin{align*}
  \overline{\rm AB}  &= 1 \\        
  \overline{\rm CD}  &= \sqrt{G^2 + r^2(1 + -\cos{2 G \rho})} \\
  \overline{\rm AC}  &= \sqrt{O^2 + r^2(1 + \sin^2{O \rho}  - \cos^2{O \rho})} \\  
  \overline{\rm AC}  &= \sqrt{O^2 + r^2(1 + -\cos{2 O  \rho})} \\
  \overline{\rm BD}  &= \sqrt{P^2 + r^2(1 + -\cos{2 P \rho}))} \\
  \overline{\rm AD}  &= \sqrt{(O+G)^2 + r^2(1 + -\cos{2 (O + G)  \rho})} \\
  \overline{\rm BC}  &= \sqrt{(P+G)^2 + r^2(1 + -\cos{2 (P + G) \rho})} \\      
\end{align*}

Using the cartesian distance formula:
\begin{align*}
  \overline{\rm CD}  &= \sqrt{G^2 + (r\sin{0 \rho} -  r\sin{G \rho})^2 + (r\cos{0 \rho} - r\cos{G \rho})^2} \\
    \overline{\rm CD}  &= \sqrt{G^2 + (0 -  r\sin{G \rho})^2 + (r - r\cos{G \rho})^2} \\
    \overline{\rm CD}  &= \sqrt{G^2 + r^2(\sin{G \rho})^2 + r^2(1 - \cos{G \rho})^2} \\
    \overline{\rm CD}  &= \sqrt{G^2 + r^2(1 + \sin^2{G \rho}  - \cos^2{G \rho})} \\
    \overline{\rm CD}  &= \sqrt{G^2 + r^2(1 + -\cos{2 G \rho})} \\        
\end{align*}
Note: $\sin^2{x \rho}  - \cos^2{x \rho} = -\cos{2 x \rho}$

So, for all O,G,P,
\begin{align*}
  \overline{\rm AB}  &= 1 \\        
  \overline{\rm CD}  &= \sqrt{G^2 + r^2(1 + -\cos{2 G \rho})} \\
  \overline{\rm AC}  &= \sqrt{O^2 + r^2(1 + \sin^2{O \rho}  - \cos^2{O \rho})} \\  
  \overline{\rm AC}  &= \sqrt{O^2 + r^2(1 + -\cos{2 O  \rho})} \\
  \overline{\rm BD}  &= \sqrt{P^2 + r^2(1 + -\cos{2 P \rho}))} \\
  \overline{\rm AD}  &= \sqrt{(O+G)^2 + r^2(1 + -\cos{2 (O + G)  \rho})} \\
  \overline{\rm BC}  &= \sqrt{(P+G)^2 + r^2(1 + -\cos{2 (P + G) \rho})} \\      
\end{align*}

>>>>

\title{Untwisting the Tetrahelix (v0.9)}
\author{Robert L. Read \texttt{read.robert@gmail.com} \and
  Robert Gatliff \texttt{robert@toubat.org}
}

@@

This formula defines a helix, but it is not any of the apparent helices, or rail helices, of the
BC helix, but rather one that winds three times as rapidly through all
nodes. To a designer of tetrahelices, it is more natural to think of
the three helices which are visually apparent, that is, those three
which are closely approximated by the outer edges or rails of
the BC helix. We think of each of these three rails as being a different color, red, blue, or yellow.

@@

Note in Figure \ref{railanglefig} the $z$-axis travel for one rail edge is denoted by $d$. In \eqref{graycoxeter} and \eqref{eq:colored}, the variable
$h$ is used for one third of the distance we name $d$. We will later justify that $d = 3h$. In this paper we assume the length of a rail
is alwyas $1$ is a simplification, although we make the rail length a parameter in our JavaScript codein \texttt{tetrahelixmath.js} \cite{readtetrahelix}.

This Gray-Coxeter formulation $V(n)$ from \eqref{graycoxeter} is valuable, but obscures the fact that we have three
apparent rail helices.
In order to rewrite this expression with an explicit rotation of $2\pi/3$, we expand 
the expressions $\theta$ and $\rho_{bc}$ and seek to isolate the term $c2\pi/3 $.
\begin{align*}
  \rho_{bc} n + c \theta  &=   \text{\{we aim for 3 in denominator, so we split...\}} \\
    (3 \theta - 2 \pi)n + (c/3)  (3 \theta)  &=   \text{\{we want $2\pi$ in numerator, so add canceling terms...\}} \\
  (3 \theta - 2 \pi)n + (c/ 3) (3 \theta - 2 \pi  + 2 \pi) &=  \text{\{associate...\}} \\
  (3 \theta - 2 \pi)n + (c/ 3) ((3 \theta - 2 \pi)  + 2 \pi) &=  \text{\{distribute...\}} \\  
  (3 \theta - 2 \pi)n + (c / 3) (3 \theta - 2 \pi)  + c 2 \pi /3 &=  \text{\{definition of $\rho_{bc}$...\}} \\
  \rho_{bc} n + (c / 3) \rho_{bc}  + c 2 \pi /3 &=  \text{\{collect like factors...\}} \\  
  n \rho_{bc}  +\rho_{bc} ( c+  2 \pi) /3  \\
\end{align*}
Now the the term on the left is the only one that depends on the scalar $n$. We use this to a create
a new formulation: $H_{BCsymmetric}(n,c) = H_{BCcolored}(n,c)$

Recall that $c \in \{0,1,2\}$, but $n$ is continuous (rational or real-valued.)

%% Actually, I am not sure if this is a clearer formulation...it might be, and need to put it in general!!!
\begin{equation}
H_{BCsymmmetric}(n,c) =
\left [
  \begin{tabular}{c}
    $ r  \cos(\rho_{bc} n + c (\rho_{bc} +  2 \pi) /3) $\\
   $ r  \sin(\rho_{bc} n + c (\rho_{bc} +  2 \pi) /3) $\\
   $ (n+c/3)  h_{bc} $
  \end{tabular}
  \right ],
\text{where:}
  \begin{tabular}{c}
%%  $\kappa = n + c/3$ \\
    $\rho_{bc} = (3 \theta - 2 \pi)$ \\
    $ \theta = \arccos(-2/3) $ \\
    $ h_{bc} = 1/\sqrt{10} $ \\    
  \end{tabular}      
\end{equation}

Since:
\[ \frac{2 \pi}{\rho_{bc}} \approx 10.16
\]
We can see that there are approximately $10.16$ red, blue or yellow tetrahedra on one rail in a single revolution.
The pitch of the Boerdijk--Coxeter helix of edge length $1$ is the length of three tetrahedra times this number:
\begin{align*}
  &= \frac{d_{bc} 2 \pi }{\rho_{bc}} \\  
%%  &= \frac{3 \cdot h_{bc} 2 \pi }{\rho_{bc}} \\
  &= \frac{3  \sqrt{\frac{2}{5}}  \pi}{\rho_{bc}} \\
  &\approx 9.64 \\
\end{align*}
The pitch is less than the number of tetrahedra because the tetrahedra
are not lined up perfectly.  It is a famous and interesting result
that the pitch is irrational. A BC helix never has two tetrahedra at
precisely the same orientation around the $z$-axis. However, this is
inconvenient to designers, who might prefer a rational pitch.
The idea of developing a rational period by arranging solid tetrahedra by relaxing the face-to-face matching
has been explored\cite{sadler2013periodic}. In this paper
we develop slightly irregular edge lengths that support, for example, a pitch of precisely 12
tetrahedra in one revolution which would allow an architect to design a
column having a basis and a capital in the same relation to the
tetrahedra they touch at the bottom and top of the column.

@@

If all three rails do not have the same pitch, there is an edge of
unbounded length, so it is hardly optimal. So we are justified in talking about the
\emph{pitch} of 
the tetrahelix as the pitch of its rail helices, even though there are
three such helices.

@@

Since the axes are parallel, we may define the \emph{inradius} of a
tetrahelix to be the radius of the largest
cylinder parallel to this axis contained within the circumscribing cylinder w
hich is penetrated by no edge.
Define an \emph{minimax edge-length optimal tetrahelix} or just an
\emph{optimal tetrahelix} to be a tetrahelix for which there exists
no other tetrahelix of the same inradius and pitch with a lower maximum
difference between its edge lengths.

@@

Since all three rails have the same rail length, no matter how we
move the rails in the $xy$ plane if we shorten the $xy$ distance between
vertices we shorten the total distance.
Our tool for thinking about this is to project out the $z$ dimension to form
a two-dimensional figure of nodes and non-rail edges.
Consider the projection along the $z$ axis of all vertices and non-rail edges into the $xy$ plane, which will be
a figure of dots and connecting segments in the $xy$ plane. The convex
hull for any one helix projection will be a circle (if its pitch is
irrational) or a polygon if rational, or a point if the helix has
zero curvature. Each of these figures by definitions lies outside the
circle of inradius $i$ in the $xy$ plane.

@@

Figure \ref{untouchingrailfig} depicts a projection of
only the point $P$, the point $P'$ which is an adjacent vertex on the same rail,
and the intermediary points $R$ and $Q$ on the other two rails.

Supoose there is a rail $P$ which does not have a segment touching the incircle
at all in the projection, as depicted in 
Then a segment connecting the other two $Q$ and $R$ is a longest
rail, because any chord touching an incircle is longer than any which does not.
These two rails ($Q$ and $R$) can be moved closer to each other, and further away from $P$, to form a better minimax solution
by shortening the longest rail, until one of the segments from $P$ the previously
untouching rail touches one of the incircle. By induction, then all rails have
at least one segement touching each vertex on the rail that is tangent to the
incircle in an optimal tetrahelix.

@@

Theorem \ref{thm:cylinder} is not true when $\rho$ is an odd integer multiple of $\pi$.

Now that we have shown that any optimal tetrahelix vertices
are on helices of the same axes and pitch, we see that the vertices 
of any optimal tetrahelix will lie on a cylinder, or a circle when axis dimension
is projected out. Therefore it is reasonable to now speak of the \emph{radius} $r$
of a tetrahelix as the radius of the cylinder, as distinct from the inradius $i$.

Now that we have coincident axes, the same pitch, and the same radius, we can go on to
 the harder proof about where vertices occur along the $z$-axis.

@@

 Note that from the point of view of a single edge, we are on a
 slanted cylinder, when $\rho \neq 0$.  This means from its point of
 view a cross section is an ellipse. So we have to be very careful in
 comparing lengths of edges relative to the tetrahedron, because a
 change in position along the edge changes the length of a line, but
 in a complicated way depending on where it is relative to the
 ellipse.

@@

\begin{theorem}
  \label{eventhirds}
  An optimal tetrahelix of any rail angle $\rho < \pi$ has all nodes evenly spaced at $d/3$ intervals on the $z$ axis.
  Any one tetrahedron in a tetrahelix has $1$ rail edge, $2$ one-hop edges connected to the rail and $2$ two-hop edges connected to the rail.
  The edge opposite of the rail edge is a one-hop edge.
\end{theorem}

@@

    Every vertex is connected to 4 non-rail edges. There is a one-hop edge
    in both the positive and negative $z$ direction. Likewise there is a two-hop
    edge in both the positive and negative $z$ direction. Let $A$ be the set
    of edge lengths, which has only 3 members, represented by $A = \{o,t,r\}$ for
    the one-hop, two-hop, and rail edge lengths.

    Any attempt to move any rail in either $z$ direction lengthens one two-hop edge to $t' > t$
    edge and shortens one one-hop edge $o' < o$. Let $B = \{o',t' \} \cup A$ be new edges.
    The minimax of $B$ > $A$ since there is a single rail length which cannot be both greater
    than $t'$ and $o'$ and less than $t'$ and $o'$.
    Therefore, any optimal tetrahelix has all one-hop edges between all rails are equal to each other, and
    all two-hop edges are equal to each other, and the $z$ distances between rails are equal, and therefore
    $d/3$ from each other.
\end{proof}

@@

@@

Rewriting our formulation in terms of $\rho$:
\begin{equation}
  \label{eq:general}
H_{general}(\chi,n,c,\rho,d_{\rho},r_{\rho}) =
\left [
  \begin{tabular}{c}
   $ r_{\rho} \cos(\chi \cdot (n \rho + c(\rho +  2 \pi) /3)) $\\
   $ r_{\rho}  \sin(\chi \cdot (n \rho + c(\rho +  2 \pi) /3)) $\\
   $ d_{\rho} (n + c /3) $
  \end{tabular}
  \right ],
\text{where:}
\begin{tabular}{c}
  $   1 = d_{\rho}^2 + 4 r_{\rho}^2 (\sin{ \rho / 2})^2 $ \\
%%  $\kappa = n + c/3$ \\
    $\chi \in \{-1,1\}$ \\  
  \end{tabular}      
\end{equation}

@@

\begin{align*}
  \text{one-hop} &= dist(H_{general}(n,c,\rho,d_{\rho}),H_{general}(n,c+1,\rho,d_{\rho}),r_{\rho})  \\
%%  \text{one-hop} &= dist(H_{general}(0,0,\rho,d_{\rho}),H_{general}(0,1,\rho,d_{\rho},r_{\rho}))  \\  
%%  \text{one-hop}  &= \sqrt{\frac{d_{\rho}^2}{9} + (r_{\rho}\sin(0) - r_{\rho}\sin(\rho/3+1\cdot\frac{2\pi}{3}))^2  +
%%    (r_{\rho}\cos(0) - r_{\rho}\cos(\rho/3 + 1\cdot\frac{2\pi}{3}))^2} \\
%%  \text{one-hop}  &= \sqrt{\frac{d_{\rho}^2}{9} + (0  - r_{\rho}\sin(\rho/3 + \frac{2\pi}{3}))^2  +
%%    (r_{\rho} - r_{\rho}\cos(\rho/3 + \frac{2\pi}{3}))^2} \\
%%  \text{one-hop}  &= \sqrt{\frac{d_{\rho}^2}{9} + r_{\rho}^2\sin^2(\rho/3 + \frac{2\pi}{3})  +
%%    r_{\rho}^2(1 - \cos(\rho/3 + \frac{2\pi}{3}))^2} \\
  %% Formula above checks at sqrt(8/27), 1, 0, and rho_bc, h_bc, r_bc.
  \text{one-hop}  &= \sqrt{\frac{d_{\rho}^2}{9} + r_{\rho}^2(\sin^2(\rho/3 + \frac{2\pi}{3})  + (1 - \cos(\rho/3 + \frac{2\pi}{3}))^2)} \\
  \text{but: }  d_{\rho}^2 &= 1 - 4 r_{\rho}^2 (\sin( \rho / 2)^2 \text{ ...so we substitute:}\\
  %% Formula above checks at sqrt(8/27), 1, 0, and rho_bc, h_bc, r_bc.  
  \text{one-hop}  &= \sqrt{\frac{1}{9}  + r_{\rho}^2(-\frac{4 (\sin^2( \rho / 2))}{9} + \sin^2(\rho/3+ \frac{2\pi}{3})  + (1 - \cos(\rho/3 + \frac{2\pi}{3}))^2)} \\
  %% Formula above checks at sqrt(8/27), 1, 0, and rho_bc, h_bc, r_bc.    
\end{align*}

@@

If you look down the axis of an optimal tetrahelix, it happens that only the one-hop edges
come closest to the axis. In other words, they define the radius of the incircle of the
projection, or the radius of a cylinder that would just fit inside the tetrahelix.
This is useful to know if the tetrahelix is used as a structure to house a pipe or
a ladder that must bear a human or to armor a conduit. The inradius $i_\rho$ of
an optimal tetrahelix is a remarkably simple function of the radius $r$ and the rail angle $\rho$:
\begin{equation}
  \label{eq:inradius}
  i_{\rho} = \frac{r \sin{(\pi - \rho)}}{6}
\end{equation}
Which can be seen from the trigonometry of a diagram of the projected one-hop edges
connecting four sequentially numbered vertices:

@@

@@

\begin{proof}
Choosing $d = 1$ and $\rho = 0$ we use Equation \eqref{eqrhoopt} to find the radius of 
optimal minimax difference.

Substituting into \eqref{eq:general}:

\begin{align*}
  \text{one-hop} &= \sqrt{\frac{1}{9} + 3r^2}\\
\end{align*}
Then:
\begin{align*}
   1  &=  \sqrt{\frac{1}{9} + 3r^2} & \text{ solved by... }\\
   r  &= \sqrt{\frac{8}{27}} &\approx 0.5443 \\
\end{align*}
\end{proof}

This radius\footnote{Another interesting but non-optimal solution is derived by setting (one-hop + two-hop)/2 = 1,
  occurs at $r = \sqrt{35}/4$ which produces
  three length classes of $11/12, 12/12, 13/12$.}
produces a two-hop rail length of $\frac{2}{\sqrt{3}}$.
The difference between this and $1$ is $\approx 15.47\% $.

\begin{figure}[H]
  \label{fig:unfolded}
     \centering
     \includegraphics[width=0.6\textwidth]{figures/EquitetrabeamUnfolded.JPG}
     \caption{Projection Diagram}
\end{figure}

@@

 Because the equitetrabeam which has a rail angle of $0$ still has
 chirality, that is, one still must decide to connect with one-hop to
 the clockwise or counter-clockwise node, it is not poosible to build
 a smooth continuum where $\rho$ transitions from positive to negative
 which remains optimal. One can use a negative $\rho$ in $H_{general}$
 but it does not produce minimax optimal tetrahelices. In other words,
 untwisting a conter-clockwise tetrahelix to rail angle $0$ and then going
even further does produce a clockwise tetrahelix, but one in which the
 one-hop and two-hop lengths in the wrong places (that is, two-hop
 becomes shorter than one-hop.) Likewise, $\rho > \rho_{bc}$ generates
 a tetrahelix, but minimax optimality is not guaranteed by $H_{general}$.

The pitch of any tetrahelix 
where $\rho \neq 0$ is:
\begin{equation}
  \label{pitcheqn}
p(\rho) = \frac{2 \pi  \cdot d}{\rho}
\end{equation}
For a fixed $z$-axis travel $d$, this is trivial.
However, if one is computing 
$z$-axis travel from \eqref{dopt} the pitch is not simple.
However, it increases monotonically and smoothly with decreasing $\rho$, so
Equation \eqref{pitcheqn} can be easily solved numerically with a Newton-Raphson
solver, as we do on our website.
For a pitch at least $ p \geq \frac{3  \sqrt{\frac{2}{5}}  \pi}{\rho_{bc}} \approx 9.64 $,
using \eqref{dopt} produces minimax optimal tetrahelices.

In this way a rail angle can be chosen for any desired (sufficiently large) pitch, yield
the optimum radius, one-hop, and two-hop lengths an engineer needs to
construct a physical structure.

Perhaps surprisingly, the optimal untwisting is accomplished only by
changing the length of the two-hop member, leaving the one-hop memever
and rail length equivalent within this continuum.\footnote{Before deriving Equation \eqref{eqrhoopt}, we created a continuum by
using a linear interpolation between the optimal radius for the
Equitetrabeam and the BC Helix. This minimax optimum of this simpler
approach was at most 1\% worse than the optimum computed by
\eqref{eqrhoopt}.}
 However, it should
be noted that an engineer or architect may also use $H_{general}$
directly and interactively, and that minimax length optimality is a
mathematic starting point rather than the final word on the beauty and utility of
physical structures. For example, a structural engineer might increase
radius in order to resist buckling. If an equitetrabeam were actually
used as a beam, an engineer might start with the optimal tetrabeam and
dilate it in one dimension to ``deepen'' the beam. Similarly, simple
changes curve the equitetrabeam into an ``arch''.

@@
